Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields

### 2020-05-29 ans_to_reviewers/answers-to-reviewers.tex

%----------------------------------------------------------------------
\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets
  \\\vspace{5mm}
 Authors response
  \\\vspace{5mm}
\small{Manuscript: AOAS1910-027}}
\author{ }

\date{\today}

\maketitle

We are grateful to the reviewers for their comments and
suggestions. They have all been considered in this revised
version of this manuscript. We have worked hard on developing both theoretical results, computational resources, and finally on the presentation in the paper. To help us on these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.

The principal revisions made to the manuscript are as follows and are done to address the main comments by Editors and Reviewers:

\begin{enumerate}
\item The motivation for excursion sets (ESs) and in particular that of {\bf{multivariate ESs}} has been strengthened. Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis. Another example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water. Our motivation in this paper is that joint temperature and salinity excursions are proxies for several processes in the oceans. For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive. Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.

In the revised manuscript these aspects are largely rewritten in the Introduction, Background, expected integrated Bernoulli variance (EIBV) and Case Study. This change addresses the comments of AE1, R1 (3,) MORE.

In the revised version we motivate more for statisticians, in particular we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.

\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of ESs and having more general results related to the EIBV for vector-valued processes, using marginal (one-location) entries and for volume calculations requiring multiple locations. The revised version highlights these novel statistical results as a main contribution of the paper.

In the revised manuscript these aspects are largely rewritten in the main body containing novel results for EIBV for sums of marginal variances as well as for more general results inluding that of volumes of random sets, while the Appendix derives the closed-form results. This change addresses the comments of E2, AE1, R2 (7), R3 (1), MORE.

\item We realized that a contribution of the current work is available code for statisticians who could try this themselves or even extend it for their own statistical cases. The computer code could of course also be useful to practitioners having related sampling design problems. The {\bf{simulation results and code}} are thoroughly re-worked. Open-source python code is now provided for all the examples used in the paper and also for simulation studies using more nuanced statistical models (non-stationary and non-isotropic).

The simulation study section is largely modified and it contains pointers to the associated open-source code. This change addresses the comments of R1 (2,3),  MORE.

@@

However, the reports also identify serious shortcomings that 
prevent me from accepting your manuscript for publication at
this time. The reports are rather consistent in the nature of 
the problems that would need to be addressed for an AOAS publication.  
Briefly, these issues include:

%\item{1. Intended audience:}\label{r3c1}
\vspace{5mm}
\noindent {\bf{Main comment 1. Intended audience:}}

As you know from reading the Manuscript Submission page of our
web site (from where you submitted your manuscript), "AOAS is 
intended to be read by all professional statisticians, as well 
as by other scientists who are interested in particular areas 
of statistical application. It is important for the abstract
and introduction to say: Why the topic is of interest, and What 
is to be shown, in as non-technical a manner as possible."
Remarks from the AE (point 1), Referee 1 (point 4), and Referee 2 
(R1R3.pdf, 1) refer to this question: "Who is the intended
audience?"  Quoting from Referee 1, "I am not clear about the 
intended audience for this paper.  Or to put it another way, 
for whom is this paper important?" The reports have some 
suggestions that may guide you in making the necessary critical 
revisions along these lines.

\reply: The comment by you, AE and reviewers made us realize that the initial submission had a motivation which was too applied. We were not selling the paper contributions in early enough, and it was not targeted sufficiently to professional statisticians. In the revised version we motivate more for statisticians.

Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians. 
(See also comment 1 in the list of main changes above.)

@@

%\item{2. Novelty:}\label{r3c2}
\vspace{5mm}
\noindent {\bf{Main comment 2. Novelty:}}

As noted in the AOAS mission statement, papers are expected
to include some degree of "methodological innovation."
In separate comments to the Editors, one reviewer noted,
"Methodological novelty is an issue with this paper.  Much of 
it is either standard Gaussian process calculations/formulae 
or sequential designs that were, in all but one case, previously 
proposed. Two things that could add a lot would be good code 
that others could run for the simulation studies like the authors' 
(if someone else were contemplating a similar experiment), and 
good code for finding the optimum design. That could lead to a 
fairly substantial impact of their work."

\reply: We realize that we did not stress the statistical novelty sufficiently in the submitted version. The main statistical novelty is in the closed-form calculation for the design criteria of the expected integrated Bernoulli variance (EIBV) associated with excursion sets (ESs) for multivariate processes. In the revised version this is extended to volumes (going beyond single location). We believe that part of the problem with the initial submission was the organization of these new results in the structure of the paper. This is hence re-structured in the revision.

Our actions are to i) extend the closed-form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed-form results to be presented in the main body, but with proofs in the new Appendix.

@@

%\item{3. Assumptions }\label{r3c3}
\vspace{5mm}
\noindent {\bf{Main comment 3. Assumptions }}

Referees 1 and 2, as well as the AE, call attention to some of
the assumptions (e.g., stationary and isotropic random fields),
and ask you to address how realistic they are.

\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious construction for the real-data case. In other related situations, for instance that of larger fjord, ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from ensembles of computer model simulations. For such cases, we agree that it would make sense to embed more nuanced models, for instance that of larger variances at certain locations and stronger correlations in certain directions, but for our field trial here we attempted to have as few parameters as possible and had no reason to assume more complicated forms.

That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-isotropic), but we do not conduct any inference for parameters in these models because it would require high resolution data from ocean models or satellites.

Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here. MORE

@@

\vspace{5mm}
\noindent {\bf{Main comment 4. Exposition}}

Both the AE and Referee 1 strongly advise shortening the first
five pages - or, at least, getting to the point (currently on p6)
much sooner.  Quoting from  Referee 1, 5: "I slogged through the 
first six [pages] before I discovered what the paper was to be 
about, on page 6."  You do not want to lose your readers, so I
advise you to engage them early on.  All four reviewers note
potential technical or typographical errors, so please proofread
carefully.

\reply: Based on the comments by yourself, AE and referees, we realize that the motivation part of the paper was too long.

Our actions include a much shorter Section 2 with background. We now aim to engage readers earlier on with the applied statistical focus and methodological contribution of our work.

@@

%\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.1.1]\label{ec1} 
\vspace{5mm}
\noindent \aecom \textbf{1}:

General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume. However, the authors never define exactly what a “boundary” is. Such a definition should be provided, as well as some discussion of why determining the location of a boundary is so important. Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par

\reply: The revised version has an entirely new Section 3 which includes general background for random sets and excursion sets. We added references, for instance \cite{adler2000excursion,Molchanov2005}.

@@

\vspace{1em}

%\item[1.1.2]\label{ec2} 
\vspace{5mm}
\noindent \aecom \textbf{2}:

Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6. I suggest that you wait to make these assumptions until you get to the simulation studies and example.\par

\reply: We agree that they are not necessary for the methodological contributions, so moved this specification to the simulation study and example. 
In the open-source that is now available, we implement non-stationary and non-isotropic covariance models. These represent situations that have relevance for our case. The code we post along with the paper facilitate testing so users can input more complex models to see how it influences designs.

@@

\vspace{1em}

%\item[1.1.3]\label{ec3} 
\vspace{5mm}
\noindent \aecom \textbf{3}:

Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable assumption? This could easily be checked by taking replicate measurements at each site.\par

\reply: The assumptions are conditionally independent measurements of salinity and temperature. We believe this is realistic as they are gathered from different sensors. We added a comment about this not necessarily being the case with other data that are extracted and processed by combining sensors. In the latent Gaussian random field, the variables are dependent.

@@

\vspace{1em}

%\item[1.1.4]\label{ec4} 
\vspace{5mm}
\noindent \aecom \textbf{4}:

General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid. In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press) that the quality of inferences for spatial fields can be improved substantially by taking some
observations very close to others.\par

\reply: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid. The sampling density is therefore higher than the grid resolution. Still, there is a limitation on restricting the sampling locations to finite number of fixed locations. In order to handle the computational load and navigation resources available on the AUV this simplification is necessary. % Comment Trygve: Should we add some more here? 
\vspace{1em}

@@

\vspace{1em}

%\item[1.1.6]\label{ec1} 
\vspace{5mm}
\noindent \aecom \textbf{6}:

Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par

\reply: We clarified the parameter specification. TO CHECK.

In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data. (Mentioned under further work in closing remarks.)

@@

%\item{Minor comments to the Author}\label{emc}

\vspace{5mm}
\noindent {\bf{Minor AE comments to the Authors}}
%\reply: 
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.2.1]\label{emc1} 
\item Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
to-last line of Page 12.

\reply: We have changed this to be consistent in the revised text.

@@

%\item[1.2.2]\label{emc2} 
\item Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
should be stated explicitly.

\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."

@@

%\item[1.2.3]\label{emc3} 
\item Page 19, fourth line from bottom: Change ``as also" to ``as well as"

\reply: We have changed this in the text - page 19.

%\item[1.2.4]\label{emc4} 
\item Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
of these two references, but actually there are no additional authors beyond the two listed.

\reply: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
\end{itemize}

@@

\reply: Thank you. We have improved and documented the code which is now open-source available.

\item{\revcom  :}\label{r1c3}

The things that stand out as novel in this paper are: the need for a se- quential design for driving an autonomous under water vehicle; a way of characterizing water plumes that are washed down from the land into the sea. Papers co-authored by one of the authors (Ginsbourger) of this paper inevitably means some overlap of ideas with the other papers, although the two step ahead greedy design algorithm seems novel. I seem to recall that Jon Lee published that algorithm albeit for determining the maximum entropy optimum by finding optimal points sequentially over space not time as in this paper. The literature on spatial design is huge and often involves spatial sequential designs. The handbook chapters by Zidek and Zimmerman (or is the other around) might be worth citing for the benefit of readers inter- ested in some of the background.

\reply: We have improved the augmented the discussion on other design criteria and why the ES related IBV is used in our case. In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe that contains a comment from Jon Lee \citep{bueso1998state}.

@@

I am not clear about the intended audience for this paper. Or to put it another way, for whom is this paper important? The case study or the
design methodology? Case studies offer concern specific events deemed of importance e.g. the much–studied O-ring failure long ago on a US rocket launch. Such a case study may then drive the innovation of statistical methods for an analysis of data that were collected because of that case’s importance. More likely the case study was intended as a tutorial on the use of a novel design methodology. In fact the Closing Remarks section points “the opportunities available real-time multivariable spatial data-gathering” that “statistician can exploit”.

\reply: TO DO

However, the user community is pretty tech-savvy nowadays, and that community might well be the primary beneficiary and more motivated than the statisticians to use the tools for sequentially sampling rivers for mapping plumes. That suggests including the relevant code for running the analyses, especially the simulation studies. An alterna- tive journal to a statistical one might be more appropriate from that perspective.

\reply: We have made an open-source repository available for the public. This includes commented code examples for exploration and analysis of the material presented in the paper. The code is optimized for speed to allow fast computation using Pytorch (citation). The code is available here: (TO DO link).

@@

\reply: TO DO

\item{\revcom :}\label{r1c6}

It is not clear why the binary classification and excursion sets are being used. Is there really a sharp boundary at the edge of a plume? Why not mapping the salinity and temperature fields instead? After all, on page 19, “achieving good predictive performance” is mentioned as a desirable feature of a good design. A lot of work has been done on that topic. And what about indicator kriging (along with logit regression) that generalizes indicator kriging that has been used to identify plumes? Modelling the p(x) is more like prohibit analysis-what is the advantage.

\reply: We added reference to \cite{adler2009random} on random sets and motivated this topic for environmental sensing. See our general comment 1 on top.

@@

\item{\revcom :}\label{r1c7}

 The sequential sampling approaches are not novel. they are mentioned as examples in the cited paper by Bect et al. What is novel is the study of how all these methods work first in simulation studies as well as well as in the case study. That work will be a valuable contribution.

\reply: We improved our statistical contribution. See our general comment 2 on top.

@@

\item{\revcom  :}\label{r1c8}

While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials. Furthermore it could cover design both for spatial mapping as well as indicator mapping. Finally model uncertainty could be included in the overall assessment of uncertainty. The more general issue here is what objective function should be optimized in the design? Are we in a multi-attribute criterion situation?

\reply: The traditional entropy calculations for this GP aim to reduce the log determinant of the updated covariance matrix at the grid locations. It will not vary with the data realizations, so in this setting this information criteria will always give the same path. The EIBV criterion is aligned with our ES motivation for this study. When focusing on the expected EIBV, the paths will vary with the data as we see in our simulation study and in the real-data example. We see this as a clear benefit of the approach compared with entropy-based calculations. Still, sampling approaches using hybrid criteria or multi-attribute criteria (Multi-objective optimization and Pareto fronts) might balance goals in such situations, but we leave that for future research (closing remarks).

@@

\item{\revcom  :}\label{r1c9}

On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic. My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase. But what I was able to show in each case is a cross-validatory assessment that showed the model was “useful”. I wonder if such a model assessment could be made here to support the assumption of separability and other things as well. By the way, in spatial mapping at least, point predictors are pretty robust against mis-specification of the covariance. Where things go sideways is in the confidence bands or ellipsoids they produce. What are the error bands when modelling binary responses? Surely the predicted plumes have fuzzy boundaries.

\reply: We focus on point-wise predictions in the criterion. Our trends are simple, so another covariance might be overfitting too much to data. We could try to fit different ranges to temperature and salinity, while maintaining separability. Could also refer to non-separable models, but pulls focus away from the current paper's focus. There are even tests for separability, but not worth so much, what do you do if you reject? Spatio-temporal variations are more interesting here, e.g. \cite{sigrist2015stochastic,richardson2017sparsity}. Ocean models, though rarely available on the scale of river plumes, might be highly useful for training - more complex trends and covariance model could then maybe be fit from the data.

@@

\reply: TO DO LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.

\item{\revcom :}\label{r1c11}

Are these results generalizable? In other words, how much plume-to-plume variability is there? One of the problems with modelling using a lot of parameters, is there instability over replicate experiments. To put it another way which model parameters represent plume-population parameters and which represent random plume effects?

\reply: This is a simple model for plume temperature and salinity variation. However, it is not common to have ocean model data or satellite data at this resolution, which would be required for training.

We have chosen to set few parameters, which also means that it is relatively easy to tune them at a break in the operation, if so is necessary.

@@

\item{\revcom :}\label{r1c12}

The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.

\reply: As pointed out above, we believe that spatio-temporal variations would be a more realistic extension, especially if repeated satellite data or ocean modeling is available. But this is likely relevant at a larger scale, such as that of fronts in a fjord or ocean interface. The methodological approach for sampling designs would however still be the same and this is the main contribution of the current paper.

@@

\item{\revcom : (specific comments)}\label{q11}

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

### 2020-05-29 main_new.tex
Revised paper and reply letter.

% "Title of the paper"
\title{Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets}
\runtitle{Excursion Probabilities for Informative Sampling}

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e1]{cedric.travelletti@stat.unibe.ch}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e4]{kanna.rajan@fe.up.pt}}

@@

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
\printead{e3}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e4}}

@@

  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management. Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provides new opportunities for experimental designs. In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena. Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework. We demonstrate how this criterion gives a description of the boundary uncertainty, that can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective. We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.

@@

Motivating examples relevant for ESs of multivariate processes are abundant. In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis. In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES. Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.

The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV; Section \ref{sec:GP_EP} sets up the required statistical modeling assumptions of the bivariate GPs for temperature and salinity; Section \ref{sec:sur} builds on these assumptions when deriving the analytical derivation of the design criteria; Section \ref{sec:heuristics} extends these results and present heuristics for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies. Section \ref{sec:case_study} demonstrates the methodology used in field work characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.

@@

The AUVs, see Fig. \ref{fig:lauv}, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF). The in-water operation time capacity depends on survey speed, payload sensors and navigation. Generally this amounts to 8-24 hrs. operation for smaller crafts, and 24-48 hrs. for larger crafts. AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.

@@

The application that is of focus in this work is towards spatial characterization of a frontal system generated by river plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity. Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects, wind, and temperature differences, this boundary often gets distorted. Initial knowledge about the location and evolution of these but the locations of these features are highly uncertain, making deterministic planning challenging.

@@

In order to formalize the problem in mathematical terms,
let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in D}$, where $D$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary. Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by 
$$
\es=\gp^{-1}(\T)=\{\x \in D: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields. 
While some aspects of the developed
approaches do not call for a specific form of $\T$, we will often stick for
simplicity to the case of orthants ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where $t_1,\dots, t_{\no} \in \R$) as this will allow efficiently calculating several key quantities. Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations. We will keep the abstract notation $\T$ and it shall be clear from the context when specific properties following the orthant assumption come into play. Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \cite{Molchanov2005}. In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.

For our application, we let $t_1$ be a threshold for temperature and $t_2$ a threshold for
salinity. The ES is then defined by $\mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}$.
This ES holds random indicator variables while the associated EPs are $p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S)$, $\bx \in \mathcal{M}$.
If the EPs are
close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean. It is, however,
challenging if the probability is close to $0.5$.

Relate to ES and refer to 
\cite{adler2000excursion,adler2009random}.

Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.

@@

%That $\mu(\es)$ is a well-defined random variable follows from
\begin{propo}
	\label{propo1}
	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
	\begin{equation*}
	\begin{split}
	\mathbb{E}[\mes(\es)^r]
	&=\int_{D^{r}} \jointExcuProb 
	\productMeasure
	,
	\end{split}
	\end{equation*}

	where the product measure is denoted as
	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
	Here $\gp$ is defined on $D$, and for
	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
	\medskip

@@

%\begin{equation}
%E(\min \{ p,(1-p)\})=\int \min P(\xi_T \leq t, \xi_S \leq s),[1-P(\xi_T \leq t, \xi_S \leq s)] p(E(\bxi)=) dE(\bxi)=,
%\end{equation}
%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.

@@

\section{Sequential updating and heuristic path planning}\label{sec:heuristics}

>>>>

% "Title of the paper"
\title{Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields}
\runtitle{Excursion Probabilities for Informative Sampling}

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e2]{cedric.travelletti@stat.unibe.ch}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e3]{jo.eidsvik@ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e4]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e5]{kanna.rajan@fe.up.pt}}

@@

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e2}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e4}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e5}}

@@

  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management. Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provides new opportunities for experimental designs. In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena. Specifically, we define a design criterion based on improved characterization of the uncertainty in the excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework. We demonstrate how this criterion can be used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective. We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.

@@

Motivating examples relevant for ESs of multivariate processes are abundant. In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis. In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \citep{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES. Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.

The remainder of this paper is organized as follows: Section \ref{sec:bg} provides select background on ocean sampling; Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for vector-valued GPs; Section \ref{sec:heuristics} builds on these assumptions when deriving the sequential design criteria for adaptive sampling; Section \ref{sec:simulations} discusses properties of the methods in simulation studies. Section \ref{sec:case_study} demonstrates the methodology used in field work characterizing a river plume; and Section \ref{sec:concl_disc} contains a summary and a discussion of future work.

@@

The AUVs considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF). The in-water operation time capacity depends on survey speed, payload sensors and navigation. Generally this amounts to 8-24 hrs. operation for smaller crafts, and 24-48 hrs. for larger crafts. AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.

@@

The application that is of focus in this work is towards spatial characterization of a frontal system generated by river plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity. Because of the local topography and the Coreolis force %\citep{coriolis1835memoire} 
the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects, wind, and temperature differences, this boundary often gets distorted. Initial knowledge about the location and evolution of these but the locations of these features are highly uncertain, making deterministic planning challenging.

@@

In order to formalize the problem in mathematical terms,
let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary. Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by 
$$
\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields.

While some aspects of the developed
approaches do not call for a specific form of $\T$, we will often stick for
simplicity to the case of orthants ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where $t_1,\dots, t_{\no} \in \R$) as this will allow efficiently calculating several key quantities. Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations. In the application specific properties following the orthant assumption come into play: We let $t_1$ be a threshold for temperature and $t_2$ a threshold for salinity. The set of index point then defines $\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2\}$. The associated EPs are $p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2]  \leq t_2)$, $\x \in \mathcal{M}$.

For statistical generality, we will however keep the abstract notation with $\T$ and general $\es$ for most of the derivations. Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \citep{Molchanov2005,adler2009random}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s. 
Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.

@@

%That $\mu(\es)$ is a well-defined random variable follows from
\begin{propo}
	\label{propo1}
	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
	\begin{equation*}
	\begin{split}
	\mathbb{E}[\mes(\es)^r]
	&=\int_{\mathcal{M}^{r}} \jointExcuProb 
	\productMeasure
	,
	\end{split}
	\end{equation*}

	where the product measure is denoted as
	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
	Here $\gp$ is defined on $\mathcal{M}$, and for
	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
	\medskip

@@

%\begin{equation}
%E(\min \{ p,(1-p)\})=\int \min P(\xi_1 \leq t, \xi_2 \leq s),[1-P(\xi_1 \leq t, \xi_2 \leq s)] p(E(\bxi)=) dE(\bxi)=,
%\end{equation}
%for the complementary probabilities, this is again evaluated by the corner regions for the block, and $\Phi_4()$ evaluations are required.

@@

\section{Sequential designs and heuristic path planning}\label{sec:heuristics}
